### Differential expression analysis for different mESC populations 


```{r}
library(tidyverse)
library(ggrepel)
library(limma)
library(ggprism)
myColors = c("#5f0f40", '#9a031e', '#f8973d', '#c05310', '#0f4c5c')
```



```{r}
sc_mesc <- read.delim('output/mesc_summary_direct.txt', sep = "\t") %>% as_tibble()
raw_quant <- read.delim('output/mesc_QCpass_quant_direct.txt', sep = "\t") %>% as_tibble()

raw_quant_long <- raw_quant %>% gather(Cell, Abundance, -PG.ProteinGroups)

raw_norm <- raw_quant %>% select(-PG.ProteinGroups)
raw_norm <- data.frame(apply(log2(raw_norm), 2, function(x) (x - median(x))/mad(x))) %>% 
            gather(Cell, Abundance)

pca_unfiltered <- read.delim('output/mesc_pca_unfiltered_direct.txt', sep = "\t")
```


```{r}

med_stats = sc_mesc %>% group_by(Plate) %>% filter(Detected_Proteins > 400) %>%
                summarise(Detected_Proteins = median(Detected_Proteins))
#Protein detection
ggplot(sc_mesc, aes(y = Detected_Proteins, fill = CellType, color = CellType)) + 
        geom_histogram(alpha= 0.5) +
        geom_hline(data=med_stats, aes(yintercept = Detected_Proteins, color = Plate), size = 1, linetype= 'dashed') +
        facet_wrap(~Plate) +
        scale_fill_manual(values=myColors) +
        scale_color_manual(values=myColors) +
        theme_prism() +
        annotate("rect", xmin = 0, xmax = 40, ymin = 0, ymax = 400,
           alpha = .4,fill = "grey20") +
        theme(legend.position ='none') +
        xlim(c(0,40))

#Protein quantification completeness 

comp <- data.frame(Accessions = raw_quant$PG.ProteinGroups, quantified = apply(raw_quant %>% select(-PG.ProteinGroups), 1, function(x) length(x) - sum(is.na(x)))) %>% 
                mutate(fraction = quantified/(dim(raw_quant)[2] - 1))

ggplot(comp, aes(x = fraction)) +
    geom_histogram(aes(y = after_stat(count / sum(count))), bins = 30, fill = myColors[5], color = myColors[5], alpha = .5) +
  scale_y_continuous(labels = scales::percent) +
  theme_prism() +
  xlab('Protein completeness') + ylab('Percentage, %')

#FACS accounting for cell size 
ggplot(sc_mesc, aes(y = FSC.A, fill = CellType, color = CellType)) + 
        geom_histogram(aes(x = ..density..), alpha= 0.5) +
        geom_density(fill=NA, size = 1) +
        facet_wrap(~Plate) +
        scale_fill_manual(values=myColors) +
        scale_color_manual(values=myColors) +
        theme_prism() +
        theme(legend.position ='none')

ggplot(sc_mesc, aes(x = FSC.A, y = DAPI.A.Compensated, color = CellType)) + 
        geom_point(alpha = 0.5, size = 2) +
        scale_color_manual(values=myColors) +
        theme_prism() +
        theme(legend.position ='top') +
        ylab('DAPI-A') + xlab('FSC-A')

ggplot(sc_mesc %>% filter(QC_Pass == 'True'), aes(x = FSC.A, y = DAPI.A.Compensated, color = Detected_Proteins)) + 
        geom_point(alpha = 0.8, size = 2) +
        theme_prism() +
        facet_wrap(~Plate) +
        theme(legend.position ='right') +
        ylab('DAPI-A') + xlab('FSC-A') +
        #viridis::scale_color_viridis(option='D')
        scale_color_gradientn(colours = rev(c("#03045e","#023e8a","#0077b6","#0096c7","#00b4d8","#48cae4","#90e0ef","#ade8f4","#caf0f8")))


#Show normalization 

ggplot(pca_unfiltered, aes(x = `PC1.27.2.`, y = `PC2.20.5.`, color = Plate)) + 
    geom_point(alpha=0.6, size = 2) +
    scale_fill_manual(values=myColors) +
    scale_color_manual(values=myColors) +
    theme_prism(border = T) +
    theme(legend.position ='right') +
    xlab('PC1 (27.2%)') + ylab('PC2 (20.5%)') +
    annotate("rect", xmin = 0.09, xmax = 0.34, ymin = -0.09, ymax = 0.1,
           alpha = .2,fill = "grey20") 

```


```{r}


#Protein detection
ggplot(sc_mesc %>% filter(QC_Pass == 'True'), aes(y = Detected_Proteins, fill = CellType, color = CellType)) + 
        geom_histogram(alpha= 0.5) +
        geom_hline(data=med_stats, aes(yintercept = Detected_Proteins, color = Plate), size = 1, linetype= 'dashed') +
        facet_wrap(~Plate) +
        scale_fill_manual(values=myColors) +
        scale_color_manual(values=myColors) +
        theme_prism() +
        theme(legend.position ='none') +
        xlim(c(0,30))

```



```{r}
# Importing data 
quant <- read.delim('output/mesc_filtered_quant_direct.txt', sep ="\t")
pca <- read.delim('output/mesc_pca_direct.txt', sep = "\t")
umap <- read.delim('output/mesc_umap_direct.txt', sep = "\t")
pmap <- read.delim('output/mesc_pmap_direct.txt', sep = "\t")


#remove keratins from data 
keratins = read.delim('data/MESC/mus_musculus_keratin.tsv', sep = "\t")
```

```{r}

### PLATE

#PCA plot
ggplot(pca, aes(x = `PC1.32.9.`, y = `PC2.24.1.`, color = Plate)) + 
   geom_point(alpha=0.6, size = 2) +
    scale_fill_manual(values=myColors) +
    scale_color_manual(values=myColors) +
    theme_prism(border = T) +
    theme(legend.position ='right') +
    xlab('PC1 (32.9%)') + ylab('PC2 (24.1%)') 

#UMAP plot
ggplot(umap, aes(x = UMAP1, y = UMAP2, color = Plate)) + 
    geom_point() +
    scale_fill_manual(values=myColors) +
    scale_color_manual(values=myColors) +
    theme_prism(border = T) +
    theme(legend.position ='right') +
    xlab('UMAP1') + ylab('UMAP2')


#Cluster
#UMAP plot
ggplot(umap, aes(x = UMAP1, y = UMAP2, color = factor(Cluster))) + 
    geom_point() +
    scale_fill_manual(values=myColors) +
    scale_color_manual(values=myColors) +
    theme_prism(border = T) +
    theme(legend.position ='right') +
    xlab('UMAP1') + ylab('UMAP2')

ggplot(pmap, aes(x = UMAP1, y = UMAP2, color = factor(Cluster))) + 
    geom_point(size = 3, alpha = .7) +
    scale_color_manual(values= myColors) +
    theme_prism(border=T)

ggplot(pmap, aes(x = UMAP1, y = UMAP2, color = factor(Plate))) + 
    geom_point(size = 3, alpha =.7) +
    scale_color_manual(values= myColors) +
    theme_prism(border=T)
```


#### Differential expression between cell types (plates)

```{r}
#Setting up design matrix for experiment
group <-  pmap$Plate
design <- model.matrix(~ 0 + group);colnames(design) <- unique(group)
design
#First 
data <- quant[-1]
data2 <- apply(log2(data),2, function(x) (x - median(x)))


contrast <- makeContrasts(m15 - m2i, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt <- topTable(fit2, sort.by = "none", number = Inf)
tt$Accessions <- quant$PG.ProteinGroups
tt <- tt %>% 
        mutate(FDR_log10 = log10(adj.P.Val) * (-1),
               Score = sqrt(FDR_log10**2 + logFC**2)) 
candidate <- tt %>% top_n(Score,n=20)
tt$candidate <- tt$Accessions %in% candidate$Accessions

ggplot(tt, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions, color = candidate)) + 
        geom_point() +
        geom_text_repel(data = tt %>% filter(-log10(adj.P.Val) > 5)) +
        geom_point(data = tt %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20') +
        scale_color_manual(values = myColors[c(2,5)])
``` 

```{r}
#Apply these proteins to the UMAP 
selected <- tt %>% filter(candidate == TRUE) %>% pull(Accessions)
prot <- quant %>% filter(PG.ProteinGroups %in% selected) %>% mutate(across(as.numeric(2:518))) %>% select(-PG.ProteinGroups)
prot <- t(prot)

rownames(prot) <- NULL
colnames(prot) <- quant %>% filter(PG.ProteinGroups %in% selected) %>% pull(PG.ProteinGroups)

prot <- t(apply(log2(prot), 2, function(x) (x - median(x))/mad(x)))
prot <- apply(prot, 1, function(x) (x - mean(x))/sd(x))

umap_prot <- cbind(umap, prot)

umap_prot <- umap_prot %>% gather('Protein', 'Abundance', 26:45)

for (i in selected){

    p <- ggplot(umap_prot %>% filter(Protein == i), aes(x = UMAP1, y = UMAP2, color = Abundance)) + 
        geom_point() +
        viridis::scale_color_viridis(option = 'H') +
        theme_prism(border = T) +
        ggtitle(paste('m15-m2i:', i))
    show(p)
}


```

```{r}
ggplot(umap_prot, aes(x = Abundance, fill= factor(Plate))) + 
    geom_histogram(position='dodge') +
    facet_wrap(~Protein, scale='free') + 
    theme_prism(border =T) +
    scale_fill_manual(values =myColors[c(2,5)])

```

```

### Differential expression between specific clusters

```{r}
#Setting up design matrix for experiment
group <- paste('Cluster', pmap$Cluster, sep = "_")
design <- model.matrix(~ 0 + group);colnames(design) <- unique(group)
design
#First 
data <- quant[-1]
data2 <- apply(log2(data),2, function(x) (x - median(x)))


## m2i clusters
contrast <- makeContrasts(Cluster_1 - Cluster_3, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt <- topTable(fit2, sort.by = "none", number = Inf)
tt$Accessions <- quant$PG.ProteinGroups
tt <- tt %>% 
        mutate(FDR_log10 = log10(adj.P.Val) * (-1),
               Score = sqrt(FDR_log10**2 + logFC**2))

candidate <- tt %>% top_n(Score,n=20)
tt$candidate <- tt$Accessions %in% candidate$Accessions

ggplot(tt, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions)) + 
        geom_point() +
        geom_text_repel(data = tt %>% filter(-log10(adj.P.Val) > 25)) +
        geom_point(data = tt %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

```

```{r}
#Apply these proteins to the UMAP 
selected <- tt %>% filter(candidate == TRUE) %>% pull(Accessions)
prot <- quant %>% filter(PG.ProteinGroups %in% selected) %>% mutate(across(as.numeric(2:518))) %>% select(-PG.ProteinGroups)
prot <- t(prot)

rownames(prot) <- NULL
colnames(prot) <- quant %>% filter(PG.ProteinGroups %in% selected) %>% pull(PG.ProteinGroups)

prot <- t(apply(log2(prot), 2, function(x) (x - median(x))/mad(x)))
prot <- apply(prot, 1, function(x) (x - mean(x))/sd(x))

umap_prot <- cbind(umap, prot)

umap_prot <- umap_prot %>% gather('Protein', 'Abundance', 26:45)

for (i in selected){

    p <- ggplot(umap_prot %>% filter(Protein == i, Cluster != 4), aes(x = UMAP1, y = UMAP2, color = Abundance)) + 
        geom_point() +
        viridis::scale_color_viridis(option = 'D') +
        theme_prism(border = T) +
        ggtitle(paste('m2i-Clusters:', i))
    show(p)
}


```

```{r}

ggplot(umap_prot, aes(x = Abundance, fill= factor(Plate))) + 
    geom_histogram(position='dodge') +
    facet_wrap(~Protein, scale='free') + 
    theme_prism(border =T) +
    scale_fill_manual(values =myColors[c(2,5)])

```


```{r}
## m15 weird clusters
#Setting up design matrix for experiment
group <- ifelse(paste('Cluster', pmap$Cluster, sep = "_") == 'Cluster_4', 'Cluster_4', 'ALL')
design <- model.matrix(~ 0 + group);colnames(design) <- unique(group)
design
#First 
data <- quant[-1]
data2 <- apply(log2(data),2, function(x) (x - median(x)))


## m2i clusters
contrast <- makeContrasts(Cluster_4 - ALL, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt <- topTable(fit2, sort.by = "none", number = Inf)
tt$Accessions <- quant$PG.ProteinGroups
tt <- tt %>% 
        mutate(FDR_log10 = log10(adj.P.Val) * (-1),
               Score = sqrt(FDR_log10**2 + logFC**2))

candidate <- tt %>% top_n(Score,n=20)
tt$candidate <- tt$Accessions %in% candidate$Accessions

ggplot(tt, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions)) + 
        geom_point() +
        geom_text_repel(data = tt %>% filter(-log10(adj.P.Val) > 25)) +
        geom_point(data = tt %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

```

```{r}
#Apply these proteins to the UMAP 
selected <- tt %>% filter(candidate == TRUE) %>% pull(Accessions)
prot <- quant %>% filter(PG.ProteinGroups %in% selected) %>% mutate(across(as.numeric(2:518))) %>% select(-PG.ProteinGroups)
prot <- t(prot)

rownames(prot) <- NULL
colnames(prot) <- quant %>% filter(PG.ProteinGroups %in% selected) %>% pull(PG.ProteinGroups)

prot <- t(apply(log2(prot), 2, function(x) (x - median(x))/mad(x)))
prot <- apply(prot, 1, function(x) (x - mean(x))/sd(x))

umap_prot <- cbind(umap, prot)

umap_prot <- umap_prot %>% gather('Protein', 'Abundance', 26:45)

for (i in selected){

    p <- ggplot(umap_prot %>% filter(Protein == i), aes(x = UMAP1, y = UMAP2, color = Abundance)) + 
        geom_point() +
        viridis::scale_color_viridis(option = 'D') +
        theme_prism(border = T) +
        ggtitle(paste('Cluster4 vs ALL:', i))
    show(p)
}


```


```{r}

contrast <- makeContrasts(Cluster_0 - Cluster_2, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt_3_0 <- topTable(fit2, sort.by = "none", number = Inf)
tt_3_0$Accessions <- quant$PG.ProteinGroups

ggplot(tt_3_0, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions)) + 
        geom_point() +
        geom_text_repel(data = tt_3_0 %>% filter(adj.P.Val < 0.1)) +
        geom_point(data = tt_3_0 %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

contrast <- makeContrasts(Cluster_0 - Cluster_1, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt_1_0 <- topTable(fit2, sort.by = "none", number = Inf)
tt_1_0$Accessions <- quant$PG.ProteinGroups


ggplot(tt_1_0, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions)) + 
        geom_point() +
        geom_text_repel(data = tt_1_0 %>% filter(adj.P.Val < 0.1)) +
        geom_point(data = tt_1_0 %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

```


```{r}
## m15 clusters
contrast <- makeContrasts(Cluster_0 - Cluster_2, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt <- topTable(fit2, sort.by = "none", number = Inf)
tt$Accessions <- quant$PG.ProteinGroups
tt <- tt %>% 
        mutate(FDR_log10 = log10(adj.P.Val) * (-1),
               Score = sqrt(FDR_log10**2 + logFC**2)) 
candidate <- tt %>% top_n(Score,n=20)
tt$candidate <- tt$Accessions %in% candidate$Accessions

ggplot(tt, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions, color = candidate)) + 
        geom_point() +
        geom_text_repel(data = tt %>% filter(-log10(adj.P.Val) > 5)) +
        geom_point(data = tt %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

```

```{r}
#Apply these proteins to the UMAP 
selected <- tt %>% filter(candidate == TRUE) %>% pull(Accessions)
prot <- quant %>% filter(PG.ProteinGroups %in% selected) %>% mutate(across(as.numeric(2:518))) %>% select(-PG.ProteinGroups)
prot <- t(prot)

rownames(prot) <- NULL
colnames(prot) <- quant %>% filter(PG.ProteinGroups %in% selected) %>% pull(PG.ProteinGroups)

prot <- t(apply(log2(prot), 2, function(x) (x - median(x))/mad(x)))
prot <- apply(prot, 1, function(x) (x - mean(x))/sd(x))

umap_prot <- cbind(umap, prot)

umap_prot <- umap_prot %>% gather('Protein', 'Abundance', 26:45)

for (i in selected){

    p <- ggplot(umap_prot %>% filter(Protein == i, Cluster %in% c(0,2)), aes(x = UMAP1, y = UMAP2, color = Abundance)) + 
        geom_point() +
        viridis::scale_color_viridis(option = 'D') +
        theme_prism(border = T) +
        ggtitle(paste('m2i-Clusters:', i))
    show(p)
}


```


```{r}

contrast <- makeContrasts(Cluster_0 - Cluster_3, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt_3_0 <- topTable(fit2, sort.by = "none", number = Inf)
tt_3_0$Accessions <- quant$PG.ProteinGroups

ggplot(tt_3_0, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions)) + 
        geom_point() +
        geom_text_repel(data = tt_3_0 %>% filter(adj.P.Val < 0.1)) +
        geom_point(data = tt_3_0 %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

contrast <- makeContrasts(Cluster_0 - Cluster_1, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt_1_0 <- topTable(fit2, sort.by = "none", number = Inf)
tt_1_0$Accessions <- quant$PG.ProteinGroups


ggplot(tt_1_0, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions)) + 
        geom_point() +
        geom_text_repel(data = tt_1_0 %>% filter(adj.P.Val < 0.1)) +
        geom_point(data = tt_1_0 %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

```


### Find the proteins that separate the three m15 clusters from PCA clusters

```{r}
select_0_3 <- tt_1_0 %>% arrange(adj.P.Val) %>% filter(Accessions %in% keratins$Entry) %>% head(20) %>% pull(Accessions)
prot_0_3 <- quant %>% filter(PG.ProteinGroups %in% select_0_3) %>% mutate(across(as.numeric(2:518))) %>% select(-PG.ProteinGroups)
prot_0_3 <- t(prot_0_3)

rownames(prot_0_3) <- NULL
colnames(prot_0_3) <- quant %>% filter(PG.ProteinGroups %in% select_0_3) %>% pull(PG.ProteinGroups)

prot_0_3 <- t(apply(log2(prot_0_3), 2, function(x) (x - median(x))/mad(x)))
prot_0_3 <- apply(prot_0_3, 1, function(x) (x - mean(x))/sd(x))

pca_0_3 <- cbind(pca, prot_0_3)

pca_0_3 <- pca_0_3 %>% gather('Protein', 'Abundance', 14:33) %>% 
            mutate(Abundance = as.numeric(Abundance))

for (i in select_0_3){

    p <- ggplot(pca_0_3 %>% filter(Protein == i), aes(x = `PC1.32.9.`, y = `PC2.24.1.`, color = Abundance)) + 
        geom_point() +
        theme_prism(border = T) +
        ggtitle(paste('0-2:', i)) +
        theme(legend.position = 'none')
    show(p)
}

```



```{r}
myColors = c("#264653","#2a9d8f","#e9c46a","#f4a261","#e76f51")

#Setting up design matrix for experiment
group <- paste('Cluster', umap$Cluster, sep = "_")
design <- model.matrix(~ 0 + group);colnames(design) <- unique(group)
design
#First 
data <- quant[-1]
data2 <- apply(log2(data),2, function(x) (x - median(x))/mad(x))

contrast <- makeContrasts(Cluster_1 - Cluster_0, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt_0_1 <- topTable(fit2, sort.by = "none", number = Inf)
tt_0_1$Accessions <- quant$PG.ProteinGroups


contrast <- makeContrasts(Cluster_2 - Cluster_0, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt_0_2 <- topTable(fit2, sort.by = "none", number = Inf)
tt_0_2$Accessions <- quant$PG.ProteinGroups


contrast <- makeContrasts(Cluster_2 - Cluster_1, levels = design)
fit <- lmFit(data2, design)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2, trend = TRUE)
tt_1_2 <- topTable(fit2, sort.by = "none", number = Inf)
tt_1_2$Accessions <- quant$PG.ProteinGroups

ggplot(tt_0_2, aes(x=logFC, y = -log10(adj.P.Val), label = Accessions)) + 
        geom_point() +
        geom_text_repel(data = tt_0_2 %>% filter(adj.P.Val < 0.1)) +
        geom_point(data = tt_0_2 %>% filter(Accessions %in% keratins$Entry), size = 3, color = 'red') + 
        theme_prism() +
        geom_hline(yintercept = -log10(0.01), linetype = "dashed", color = 'grey20')

```


### Find the proteins that separate the three m15 clusters

```{r}
select_0_3 <- tt_0_2 %>% arrange(adj.P.Val) %>% filter(Accessions %in% keratins$Entry) %>% head(20) %>% pull(Accessions)
prot_0_3 <- quant %>% filter(PG.ProteinGroups %in% select_0_3) %>% mutate(across(as.numeric(2:518))) %>% select(-PG.ProteinGroups)
prot_0_3 <- t(prot_0_3)

rownames(prot_0_3) <- NULL
colnames(prot_0_3) <- quant %>% filter(PG.ProteinGroups %in% select_0_3) %>% pull(PG.ProteinGroups)

prot_0_3 <- t(apply(log2(prot_0_3), 2, function(x) (x - median(x))/mad(x)))
prot_0_3 <- apply(prot_0_3, 1, function(x) (x - mean(x))/sd(x))

umap_0_3 <- cbind(umap, prot_0_3)

umap_0_3 <- umap_0_3 %>% gather('Protein', 'Abundance', 14:33)

for (i in select_0_3){

    p <- ggplot(umap_0_3 %>% filter(Protein == i), aes(x = UMAP1, y = UMAP2, color = Abundance)) + 
        geom_point() +
        viridis::scale_color_viridis(option = 'H') +
        theme_prism(border = T) +
        ggtitle(paste('0-2:', i))
    show(p)
}

```