# Visualization of identification summaries for single-cell runs

### This script is used to plot the data from summary statistics of SN16 searches 
```{r}
library(tidyverse)
library(ggprism)
myColors2 = c("#5f0f40", '#9a031e', '#fb8b24', '#e36414', '#0f4c5c')
source('helperFunctions.R')
```


```{r}
#Importing data 

#Resaved the file with numbers to remove the thousand separator but that caused extra issues. Have to adjust the generateSNstats.py function to avoid the format issues
sc_set <- read.csv("data/SN17/SN17-R11_stats.csv", sep = "\t", dec = ",") %>% as_tibble() %>% 
                mutate(Precursors = as.numeric(gsub("\\.", "", Precursors)),
                        Peptides = as.numeric(gsub("\\.", "", Peptides)), 
                        Protein.Groups = as.numeric(gsub("\\.", "", Protein.Groups))) %>% 
             mutate(gradient = factor(str_match(X, pattern = "precol_(.+?)_240k")[,2]),
                    usedQBL = ifelse(grepl("QBL|QFL", X), "+", "-")) %>% 
             group_by(gradient, usedQBL)

sc_set_summary <- calculate_summary(sc_set)
```

## Single-cell dataset runs 

```{r}
ylimits = c(0,7000)
#Converting the dataframes to long format
sc_set_long <- sc_set %>% gather(key = 'Values', 'Counts',  Precursors:Protein.Groups) %>% 
                filter(Values != "Precursors",) %>% 
                mutate(gradient = gsub("min", "",gradient))
sc_set_summary_long <- sc_set_summary %>% gather(key = 'Values', 'Counts',  Precursors:Protein.Groups) %>% 
                filter(Values != "Precursors") %>% 
                mutate(gradient = gsub("min", "",gradient))

#Generating the plot 
ylimits = c(0,7000)
ggplot(sc_set_long %>% filter(Values == "Peptides"), aes(x = usedQBL, y = Counts, fill = gradient, color = gradient, label = Counts)) + 
    geom_bar(data = sc_set_summary_long %>% filter(Values == "Peptides"), stat = "identity", position= "dodge", alpha =0.6, show.legend = FALSE) +
    geom_jitter(width=0.2, height = 0, size = 4, alpha =.7) + 
    geom_text(data = sc_set_summary_long %>% filter(Values == "Peptides"), color = "black", vjust = -5,  fontface = "bold")+
    scale_fill_manual(values = c('#003049', '#d62828')) +
    scale_color_manual(values = c('#003049', '#d62828')) +
    facet_grid(col = vars(gradient)) +
    theme_prism()+
    scale_y_continuous(guide = "prism_offset", expand = c(0,0),limits = ylimits) +
    theme(legend.position =  "none", legend.title = , axis.ticks.x = element_blank(), 
          axis.text.x = element_text(angle = 0), panel.spacing = unit(1, "lines")) + 
    ylab('Peptides') + xlab("")

ylimits = c(0,2500)
ggplot(sc_set_long %>% filter(Values == "Protein.Groups"), aes(x = usedQBL, y = Counts, fill = gradient, color = gradient, label = Counts)) + 
    geom_bar(data = sc_set_summary_long %>% filter(Values == "Protein.Groups"), stat = "identity", position= "dodge", alpha =0.6, show.legend = FALSE) +
    geom_jitter(width=0.2, height = 0, size = 4, alpha =.7) + 
    geom_text(data = sc_set_summary_long %>% filter(Values == "Protein.Groups"), color = "black", vjust = -4.5,  fontface = "bold")+
    scale_fill_manual(values = c('#023f60', '#9c1d1d')) +
    scale_color_manual(values = c('#023f60', '#9c1d1d')) +
    facet_grid(col = vars(gradient)) +
    theme_prism()+
    scale_y_continuous(guide = "prism_offset", expand = c(0,0),limits = ylimits) +
    theme(legend.position =  "none", legend.title = , axis.ticks.x = element_blank(), 
          axis.text.x = element_text(angle = 0), panel.spacing = unit(1, "lines")) + 
    ylab('Protein groups') + xlab("")
```

# Precision analysis directDIA vs QBL 
```{r}
min29_direct <- read.csv('data/SN17/SN17-R11-29min-1cell_PG.tsv', dec = ",", sep= "\t") %>% as_tibble()
min29_qbl <- read.csv('data/SN17/SN17-R11-29min-1cell-QBL_PG.tsv', dec = ",", sep= "\t") %>% as_tibble()
min52_direct <- read.csv('data/SN17/SN17-R11-52min-1cell_PG.tsv', dec = ",", sep= "\t") %>% as_tibble()
min52_qbl <- read.csv('data/SN17/SN17-R11-52min-1cell-QBL_PG.tsv', dec = ",", sep= "\t") %>% as_tibble()

#Chaning column names 

colnames(min29_direct) <- c("Accessions", 'Genes', paste('Cell', 1:62, sep = ''))
colnames(min29_qbl) <- c("Accessions", 'Genes', paste('Cell', 1:62, sep = ''))
colnames(min52_direct) <- c("Accessions", 'Genes', paste('Cell', 1:40, sep = ''))
colnames(min52_qbl) <- c("Accessions", 'Genes', paste('Cell', 1:40, sep = ''))

#Changing the word 'Filtered' to NA and drop cells with low protein id numbers
min29_direct[min29_direct == "Filtered"] <- NA; min29_direct <- min29_direct[, colSums(is.na(min29_direct)) < 1500]
min29_qbl[min29_qbl == "Filtered"] <- NA; min29_qbl <- min29_qbl[, colSums(is.na(min29_qbl)) < 1500]
min52_direct[min52_direct == "Filtered"] <- NA; min52_direct <- min52_direct[, colSums(is.na(min52_direct)) < 1500]
min52_qbl[min52_qbl == "Filtered"] <- NA; min52_qbl <- min52_qbl[, colSums(is.na(min52_qbl)) < 1500]
```

```{r}
#Extrac proteins calculated in both methods 
in29 <- intersect(min29_direct$Accessions, min29_qbl$Accessions)
in52 <- intersect(min52_direct$Accessions, min52_qbl$Accessions)
```

### CV analysis 
```{r}
min29_cvs <- min29_qbl %>% gather(Cell, Abundance, Cell2:Cell62) %>% group_by(Accessions, Genes) %>%
                        mutate(Abundance = as.numeric(sub(",", ".", Abundance, fixed = TRUE))) %>% 
                        summarise(avg = mean(Abundance, na.rm =TRUE),
                                  std = sd(Abundance, na.rm = TRUE),
                                  cv = std/avg) %>% 
                        mutate(QBLonly = ifelse(Accessions %in% in29, "Direct", 'GPF'))

min29_cvs <- bind_rows(min29_cvs, min29_cvs %>% mutate(QBLonly = "Total")) %>%
                  mutate(QBLonly = factor(QBLonly, levels = c("Total", 'Direct', 'GPF')))

min52_cvs <- min52_qbl %>% gather(Cell, Abundance, Cell1:Cell40) %>% group_by(Accessions, Genes) %>%
                        mutate(Abundance = as.numeric(sub(",", ".", Abundance, fixed = TRUE))) %>% 
                        summarise(avg = mean(Abundance, na.rm =TRUE),
                                  std = sd(Abundance, na.rm = TRUE),
                                  cv = std/avg)  %>% 
                        mutate(QBLonly = ifelse(Accessions %in% in52, "Direct", 'GPF'))

min52_cvs <- bind_rows(min52_cvs, min52_cvs %>% mutate(QBLonly = "Total")) %>%
                  mutate(QBLonly = factor(QBLonly, levels = c("Total", 'Direct', 'GPF')))

```

### CV histograms 
```{r}
ggplot(min29_cvs, 
        aes(x = cv, color = QBLonly, fill = QBLonly)) + 
        geom_histogram(aes(y=..density..),bins = 50, alpha = .8) +
        geom_density(alpha = 0.1,size = 1) +
        facet_grid(vars(QBLonly)) + 
        scale_fill_manual(values = c('grey50','#003049', '#2f7ca6')) +
        scale_color_manual(values = c('grey50','#003049', '#2f7ca6')) +
        theme_prism() +
        theme(legend.position = "none") +
        xlab('CV') + ylab("Density") +
        xlim(c(0,1)) + ylim(c(0,5)) +
        geom_vline(xintercept = 0.1, linetype ='dashed', size = 0.8, color = 'grey') +
        geom_vline(xintercept = 0.4, linetype ='dashed', size = 0.8, color = 'grey') 

ggplot(min52_cvs, 
        aes(x = cv, color = QBLonly, fill = QBLonly)) + 
        geom_histogram(aes(y=..density..),bins = 50, alpha = .8) +
        geom_density(alpha = 0.1,size = 1) +
        facet_grid(vars(QBLonly)) + 
        scale_fill_manual(values = c('grey50','#d62828', '#a85353')) +
        scale_color_manual(values = c('grey50','#d62828', '#a85353')) +
        theme_prism() +
        theme(legend.position = "none") +
        xlab('CV') + ylab("Density") +
        xlim(c(0,1)) + ylim(c(0,5)) +
        geom_vline(xintercept = 0.1, linetype ='dashed', size = 0.8, color = 'grey') +
        geom_vline(xintercept = 0.4, linetype ='dashed', size = 0.8, color = 'grey') 

```

### CV vs abundance hexbins
```{r}
ggplot(min29_cvs %>% filter(QBLonly != 'Total'),
        aes(y = log2(avg), x= cv)) +
        geom_hex(bins=35) +
        xlim(c(0,0.3)) +
        facet_wrap(~QBLonly) +
        theme_prism() +
        #viridis::scale_fill_viridis(option = "A") +
        scale_fill_gradientn(colours = c('#bfd9e6','#0175b4')) +
        ylab("log2(Abundance)") + xlab("CV")


ggplot(min52_cvs %>% filter(QBLonly != 'Total'),
        aes(y = log2(avg), x= cv)) +
        geom_hex(bins=35) +
        xlim(c(0,0.3)) +
        facet_wrap(~QBLonly) +
        theme_prism() +
        #viridis::scale_fill_viridis(option = "A") +
        scale_fill_gradientn(colours = c('#fcd0d0','#d10303')) +
        ylab("log2(Abundance)") + xlab("CV")

```

### Calculating percentage of missing values in the datasets
```{r}
# Percentage = proteins in sample / total proteins in dataset 

mis29_direct = (dim(min29_direct)[1] - colSums(is.na(min29_direct))[3:dim(min29_direct)[2]])/dim(min29_direct)[1]
mis29_qbl = (dim(min29_qbl)[1] - colSums(is.na(min29_qbl))[3:dim(min29_qbl)[2]])/dim(min29_qbl)[1]
mis52_direct = (dim(min52_direct)[1] - colSums(is.na(min52_direct))[3:dim(min52_direct)[2]])/dim(min52_direct)[1]
mis52_qbl = (dim(min52_qbl)[1] - colSums(is.na(min52_qbl))[3:dim(min52_qbl)[2]])/dim(min52_qbl)[1]

#Combine into dataframe for plotting

comp_df <- data.frame(completeness = c(mis29_direct, mis29_qbl, mis52_direct, mis52_qbl),
                      gradient = c(rep('29min', length(mis29_direct)*2), rep('52min', length(mis52_direct)*2)),
                      QBL = c(rep('Direct', length(mis29_direct)), rep('GPF', length(mis29_qbl)),
                              rep('Direct', length(mis52_direct)), rep('GPF', length(mis52_qbl))))

comp_labels <- data.frame(gradient = c("29min", "29min", "52min", "52min"),
                          QBL = c("Direct", "GPF", "Direct", "GPF"),
                          x = 0.63, y = 15, pg = c(dim(min29_direct)[1],dim(min29_qbl)[1], dim(min52_direct)[1],dim(min52_qbl)[1]))
```

```{r}
ggplot(comp_df, aes(x = completeness, color = gradient, fill = gradient)) + 
        geom_histogram(aes(y=..density..),bins = 50, alpha = .8) +
        geom_density(alpha = 0.1,size = 1) +
        facet_grid(cols = vars(QBL), rows = vars(gradient)) +
        scale_fill_manual(values = c('#003049', '#d62828')) +
        scale_color_manual(values = c('#003049', '#d62828')) +
        theme_prism() +
        theme(legend.position = "none") +
        xlab('Data completeness') + ylab("Density") +
        xlim(c(0.55,1)) + ylim(c(0,22)) + 
        geom_vline(xintercept = 0.7, linetype ='dashed', size = 0.8, color = 'grey50') +
        geom_text(data =comp_labels, aes(x=x, y =y, label = pg), fontface = "bold")
```

```{r}
ggplot(comp_df, aes(x = completeness, color = gradient, fill = gradient)) + 
  stat_ecdf(geom = "step", size = 2) +
  facet_grid(cols = vars(QBL), rows = vars(gradient)) +
  scale_fill_manual(values = c('#003049', '#d62828')) +
  scale_color_manual(values = c('#003049', '#d62828')) +
  theme_prism() +
  theme(legend.position = "none") +
  xlab('Data completeness') + ylab("Density") +
  xlim(c(0.6,1)) +
  geom_vline(xintercept = 0.7, linetype ='dashed', size = 0.8, color = 'grey50')

```

## Importing PCA and UMAP results from python for plotting 
```{r}
sc_pca <- read.csv("results/pca.csv", sep = "\t") %>% as_tibble() 
sc_umap <- read.csv("results/umap.csv", sep = "\t") %>% as_tibble() %>% mutate(Gradient = c(rep('29min',60), rep('52min',38)))
```

```{r}
#Plotting the PCA 

ggplot(sc_pca, aes(x = `PC1.63.3.`, y = `PC2.5.2.`, color = MKI67)) + 
      geom_point(size = 3, alpha = 0.8) +
      theme_prism(border = TRUE, base_rect_size = 2) +
      xlab('PC1 (63.3%)') + ylab('PC2 (5.2%)') +
      #viridis::scale_color_viridis(option = "E")
      #scale_color_gradientn(colours = myColors2[c(1,3,5)])
      scale_color_gradientn(colours = c('#d9d9d9','#f1d5d5','#b60202'))

ggplot(sc_pca, aes(x = `PC1.63.3.`, y = `PC2.5.2.`, color = Gradient)) + 
      geom_point(size = 3, alpha = 0.8) +
      theme_prism(border = TRUE, base_rect_size = 2) +
      xlab('PC1 (63.3%)') + ylab('PC2 (5.2%)') +
      #viridis::scale_color_viridis(option = "E")
      scale_color_manual(values = myColors2[c(1,3,5)])
      #scale_color_gradientn(colours = c('#d9d9d9','#f1d5d5','#b60202'))

ggplot(sc_pca, aes(x = `PC1.63.3.`, y = `PC2.5.2.`, color = Sample)) + 
      geom_point(size = 3, alpha = 0.8) +
      theme_prism(border = TRUE, base_rect_size = 2) +
      xlab('PC1 (63.3%)') + ylab('PC2 (5.2%)') +
      viridis::scale_color_viridis(option = "E",discrete = TRUE) +
      theme(legend.position = 'none')

```

```{r}
#UMAP plot
ggplot(sc_umap, aes(x = UMAP1, y = UMAP2, color = MKI67)) + 
      geom_point(size = 3, alpha = 0.8) +
      theme_prism(border = TRUE, base_rect_size = 2) +
      xlab('UMAP1') + ylab('UMAP2') +
      #scale_color_manual(values = myColors2)
      #viridis::scale_color_viridis(option = "E")
      #scale_color_gradientn(colours = c('#d9d9d9','#f1d5d5','#7e0202'))
      scale_color_gradientn(colours = c('#e0e0e0','#f1e6df', '#ff6200'))


ggplot(sc_umap, aes(x = UMAP1, y = UMAP2, color = Gradient)) + 
      geom_point(size = 3, alpha = 0.8) +
      theme_prism(border = TRUE, base_rect_size = 2) +
      xlab('UMAP1') + ylab('UMAP2') +
      scale_color_manual(values = myColors2[c(1,3,5)])

ggplot(sc_umap, aes(x = UMAP1, y = UMAP2, color = Sample)) + 
      geom_point(size = 3, alpha = 0.8) +
      theme_prism(border = TRUE, base_rect_size = 2) +
      xlab('UMAP1') + ylab('UMAP2') +
      viridis::scale_color_viridis(option = "E",discrete = TRUE) 
```
